Constant Error Rate Computational Fuzzy Extractors

### 2014-01-10 trunk/constantRate/ConstCompFuzzExt.tex
Outline of proof (still need to fill in a lot)

\begin{lemma}
\label{lem:code bits statistically indist}
Let $X_i = W_i$ if $c_i = 1$, otherwise let $X_i = U$.  Then, for every $i$, $\Delta(View(A^{\mathcal{O}_{(X_1,..., X_\ell)}(\cdot, \cdot)})| c_i = 1, View(A^{\mathcal{O}_{(X_1,..., X_\ell)}(\cdot, \cdot)} | c_i =0) \leq \ngl(n)$.
\end{lemma}
\begin{proof}
\bnote{Do this}
\end{proof}
We can extend this lemma to the entire codeword $C$:
\begin{lemma}
For all $c_1, c_2 \in C$ we have the following: 
\[
\Delta(View(A^{\mathcal{O}_{(X_{c_{1_1}},..., X_{c_{2_\ell}})}(\cdot, \cdot)})| C = c_1, View(A^{\mathcal{O}_{(X_{c_{2_1}},..., X_{c_{2_\ell}})}(\cdot, \cdot)} | C = c_2)) \leq \ngl(n).
\]
\end{lemma}
\begin{lemma}
If $C$ is output then, for all $i$, $\Hav(C_i |View(A^{\mathcal{O}_{(X_1,..., X_\ell)}(\cdot, \cdot)}))\geq 1 - \ngl(n)$.
\end{lemma}
\begin{proof}
Since $C$ is output equiprobable we know that for each $i$, $\Pr[C_i =1 ] = 1/2$.  Let $e$ be the exponent of the particular negligible value above~(note $e = \omega(1)$).  For $b\in\{0,1\}$, by \lemref{lem:code bits statistically indist}, we have that in expectation each for each $i$, $\Pr[C_i =1 ] = 1/2$. 
\begin{align*}
\expe_{view \leftarrow View(A^{\mathcal{O}_{(X_1,..., X_\ell)}})}\Pr[C_i =b | view ] &\leq 1/2 + \ngl(n)\\
\end{align*}
If this was not true~(that is in expectation, there was some $b$ that was likely with probability greater than $1/2+1/\poly(n)$), there there would exist an unbounded adversary outputting that $b$, contradicting \lemref{lem:code bits statistically indist}).  
Taking the negative log of each side one has:
\begin{align*}
\Hav(C_i |  View(A^{\mathcal{O}_{(X_1,..., X_\ell)}(\cdot)}))
 &=\\-\log \left(\expe_{view \leftarrow View(A^{\mathcal{O}_{(X_1,..., X_\ell)}})}\max_{b\in\{0,1\}}\Pr[C_i =b | view ]  \right) &\geq -\log (1/2 + \ngl(n))\\
&\geq -\log(\frac{1}{2} + \frac{1}{n^{w}} )= -\log \left(\frac{n^{w}+2}{2n^{w}}\right)\\
%&=-\left(\log (n^w+2) -\log 2 -\log n^w \right)\\
&=1-\left(\log (1+\frac{2}{n^e}) \right)\\
&=1-\frac{1}{\ln 2}\left(\ln (1+\frac{2}{n^e}) \right)\\
&= 1- \frac{1}{\ln 2}\left(\sum_{i=1}^\infty \frac{(-1)^{i+1}}{i} (\frac{2}{n^e})^i\right) \\
&\geq 1 - \frac{1}{\ln 2} \frac{2}{n^e} = 1-\ngl(n)
\end{align*}
Here the second to last line is derived using the Taylor expansion for $\ln(1+x)$ and the last line is derived by noting it is a geometric series and thus all terms $i=3,....$ are contained in the second term $i=2$ which is positive.
\end{proof}
By similar reasoning we can extend this lemma to the entire codeword:
\begin{lemma}
If $C$ is output equiprobable then, $\Hav(C | View(A^{\mathcal{O}_{(X_{c_{2_1}},..., X_{c_{2_\ell}})}})) \geq |C| - \ngl(n)$.
\end{lemma}
%\begin{lemma}
%$\Hoo(C | View (A^{\mathcal{O}_{(X_1,..., X_\ell)}(\cdot, \cdot)})) \geq \log |C| - o(\ell \log n)$.
%\end{lemma}
\begin{corollary}
\consref{cons:first construction} is a $XXXXX$ computational fuzzy extractor.
\end{corollary}
\begin{proof}
Result of security of point obfuscation and \lemref{lem:cond and ext}.
\end{proof}

\subsection{Correctness of \consref{cons:first construction}}
Most of the effort in showing the correctness of \consref{cons:first construction} is showing that $\dis(w, w')\leq t$ implies $\dis(c, c')\leq t$.  However, we start by justifying our use of unidirectional codes.
\begin{lemma}
\label{lem:no 1 to 0 flips}
The probability of a single $1\rightarrow 0$ bit flip~(an obfuscation of a random block being interpreted as the obfuscation of the point) is $\ngl(n) \leq \ell/2^n$.
\end{lemma}
\begin{proof}
\bnote{we need to come back to this if we use obfuscation with error}
Recall that $c$ contains at most $\ell$ $1$s.  Since $w'$ is chosen independently of the points $r_i$, this probability is the size of each block, that is $\Pr[r_i =w_i']  = 1/2^n$. Since the $r_i$ are chosen independently, one has,
\begin{align*}
\Pr[\text{no $1\rightarrow 0$ flips}] &\geq 1-\Pr[\text{some }r_i = w_i']\\
&\geq 1-\ell \frac{1}{2^n} = 1-\frac{\poly(n)}{2^n} = \ngl(n)
\end{align*}
\end{proof}
\bnote{even if all $t$ are inserted into the bad block its okay right now.  My goal is to decrease the error tolerance of the code below $t$.}
We now turn to the question of how many $0\rightarrow 1$ bits flips occur between $c, c'$.  Recall that $\dis(w, w')\leq t$ furthermore recall that $w, w'$ are selected independently of the fuzzy extractor, in particular, independently of $c$.  We break the indices of $c = c_1,..., c_\ell$ into two components those where $c_i=1$ and those $c_i = 0$.  Denote these two components by $c_{I_1}$ and $c_{I_0}$.  Our goal is to bound how many of the places where $w, w'$ are inserted into $c_{I_0}$.  Indices $i$,  where $w_i \neq w_i'$ and $c_i = 1$ will be ``corrected'' as $w_i'$ yield $c_i' = 1$ with overwhelming probably~(see \lemref{lem:no 1 to 0 flips}).  Since $w_{err}$ be set of indices where $w, w'$ differ then for all $j\in w_{err}, \Pr[j\in c_{I_0}] = 1/2$.  This is a binomial distribution with $p=1/2$ and $n=t$.  Denote by $X =|\{j | w_j \neq w_j' \wedge j\in c_{I_1}|$. Thus, $\expe[X] = t/2$.  Using Hoeffding's inequality~\cite{hoeffding1963probability}, one has that 
\begin{align*}
\Pr[X\leq t(1/2+\alpha)] \leq 1-e^{-2\alpha^2 t}
\end{align*}
This leads us to the following lemma:
\begin{lemma}
\label{lem:correctness holds}
If $t = \omega(\log n)$ and that $\dis(w, w')\leq t$ then for any constant $\alpha>0$, then $\Pr[\dis(c, c')\geq t(1/2+\alpha)] \leq 1-e^{-2\alpha t} = 1-\ngl(n)$.
\end{lemma}
\textbf{Note:} If $t = O(\log n)$ the only error tolerance for $C$ that results in decoding with overwhelming probability is $t$.  However, in most use cases, we expect $t=\omega(\log n)$.

\section{Supported Sources}
We now describe a set of high entropy sources appropriate for our construction.  After giving the description of the distribution, we give examples of distributions that fit in this class and provide additional characterizations.

\begin{definition}
Let $\mathcal{O}_{w_1,..., w_k}$ be an oracle that return $\mathcal{O}_{w_1,..., w_k}(i, w_i')=\left( w_i\overset{?}=w_i'\right)$.
A source $W = W_1||...|W_k$ is a $(q, \alpha, \ell)$-\emph{guessable block distribution} if for any adversary $A$~(not bounded in time or space) with oracle access to $\mathcal{O}$ making at most $q$ queries there exists a set $S$ of size at least $\ell$ such that 
\[
\forall i\in S, \Hav(W_i |View(A^{\mathcal{O}_{W}(\cdot, \cdot)}))\geq \alpha.
\]
\end{definition}
This type of source seems to be inherently adaptive and is difficult to formulate in a clean information-theoretic notion.  However, we can show necessary and sufficient types of sources for a guessable block distribution.  We begin by defining the entropy jump of a block source which will appear in our characterization.

@@

\begin{claim}
Let $W$ be a $(q, \alpha, \ell)$-guessable block distribution.  Then for all orderings $i_1, ...., i_k$ there is some $j\leq k-\ell$ such that $J(i_j)> \log q + \alpha$.
\end{claim}
\begin{proof}
Suppose not, for convenience assume the ordering that violates this condition is $1,..., k$.  We describe a distribution $W$ where $\forall 1\leq j\leq k-\ell$, $J(j)\leq \log q +\alpha$. Define $W$ as follows $W_1 = c_1,...., W_{j-1} = c_{j-1}$ are constants.
\end{proof}
\bibliographystyle{alpha}
\bibliography{crypto}

>>>>

\begin{lemma}
\label{lem:code bits statistically indist}
Let $X_i = W_i$ if $c_i = 1$, otherwise let $X_i = U$.  Then, for every $i$, $\Delta(View(A^{\mathcal{O}_{(X_1,..., X_\ell)}(\cdot, \cdot)})| c_i = 1, View(A^{\mathcal{O}_{(X_1,..., X_\ell)}(\cdot, \cdot)} | c_i =0) \leq \ngl(n)$.
\end{lemma}
\begin{proof}
\bnote{Do this}
\end{proof}
We can extend this lemma to the entire codeword $C$:
\begin{lemma}
For all $c_1, c_2 \in C$ we have the following: 
\[
\Delta(View(A^{\mathcal{O}_{(X_{c_{1_1}},..., X_{c_{2_\ell}})}(\cdot, \cdot)})| C = c_1, View(A^{\mathcal{O}_{(X_{c_{2_1}},..., X_{c_{2_\ell}})}(\cdot, \cdot)} | C = c_2)) \leq \ngl(n).
\]
\end{lemma}
\begin{lemma}
If $C$ is output then, for all $i$, $\Hav(C_i |View(A^{\mathcal{O}_{(X_1,..., X_\ell)}(\cdot, \cdot)}))\geq 1 - \ngl(n)$.
\end{lemma}
\begin{proof}
Since $C$ is output equiprobable we know that for each $i$, $\Pr[C_i =1 ] = 1/2$.  Let $e$ be the exponent of the particular negligible value above~(note $e = \omega(1)$).  For $b\in\{0,1\}$, by \lemref{lem:code bits statistically indist}, we have that in expectation each for each $i$, $\Pr[C_i =1 ] = 1/2$. 
\begin{align*}
\expe_{view \leftarrow View(A^{\mathcal{O}_{(X_1,..., X_\ell)}})}\Pr[C_i =b | view ] &\leq 1/2 + \ngl(n)\\
\end{align*}
If this was not true~(that is in expectation, there was some $b$ that was likely with probability greater than $1/2+1/\poly(n)$), there there would exist an unbounded adversary outputting that $b$, contradicting \lemref{lem:code bits statistically indist}).  
Taking the negative log of each side one has:
\begin{align*}
\Hav(C_i |  View(A^{\mathcal{O}_{(X_1,..., X_\ell)}(\cdot)}))
 &=\\-\log \left(\expe_{view \leftarrow View(A^{\mathcal{O}_{(X_1,..., X_\ell)}})}\max_{b\in\{0,1\}}\Pr[C_i =b | view ]  \right) &\geq -\log (1/2 + \ngl(n))\\
&\geq -\log(\frac{1}{2} + \frac{1}{n^{w}} )= -\log \left(\frac{n^{w}+2}{2n^{w}}\right)\\
%&=-\left(\log (n^w+2) -\log 2 -\log n^w \right)\\
&=1-\left(\log (1+\frac{2}{n^e}) \right)\\
&=1-\frac{1}{\ln 2}\left(\ln (1+\frac{2}{n^e}) \right)\\
&= 1- \frac{1}{\ln 2}\left(\sum_{i=1}^\infty \frac{(-1)^{i+1}}{i} (\frac{2}{n^e})^i\right) \\
&\geq 1 - \frac{1}{\ln 2} \frac{2}{n^e} = 1-\ngl(n)
\end{align*}
Here the second to last line is derived using the Taylor expansion for $\ln(1+x)$ and the last line is derived by noting it is a geometric series and thus all terms $i=3,....$ are contained in the second term $i=2$ which is positive.
\end{proof}
By similar reasoning we can extend this lemma to the entire codeword:
\begin{lemma}
If $C$ is output equiprobable then, $\Hav(C | View(A^{\mathcal{O}_{(X_{c_{2_1}},..., X_{c_{2_\ell}})}})) \geq |C| - \ngl(n)$.
\end{lemma}
%\begin{lemma}
%$\Hoo(C | View (A^{\mathcal{O}_{(X_1,..., X_\ell)}(\cdot, \cdot)})) \geq \log |C| - o(\ell \log n)$.
%\end{lemma}
\begin{corollary}
\consref{cons:first construction} is a $XXXXX$ computational fuzzy extractor when $W = W_1,..., W_n$ all $W_i$ are independent and $\Hoo(W_i) = \omega(\log n)$.
\end{corollary}
\begin{proof}
Result of security of point obfuscation and \lemref{lem:cond and ext}.
\end{proof}

\subsection{Correctness of \consref{cons:first construction}}
Most of the effort in showing the correctness of \consref{cons:first construction} is showing that $\dis(w, w')\leq t$ implies $\dis(c, c')\leq t$.  However, we start by justifying our use of unidirectional codes.
\begin{lemma}
\label{lem:no 1 to 0 flips}
The probability of at least one $1\rightarrow 0$ bit flip~(an obfuscation of a random block being interpreted as the obfuscation of the point) is $ \ell/2^n = \ngl(n)$.
\end{lemma}
\begin{proof}
\bnote{we need to come back to this if we use obfuscation with error}
Recall that $c$ contains at most $\ell$ $1$s.  Since $w'$ is chosen independently of the points $r_i$, this probability is the size of each block, that is $\Pr[r_i =w_i']  = 1/2^n$. Since the $r_i$ are chosen independently, one has,
\begin{align*}
\Pr[\text{no $1\rightarrow 0$ flips}] &\geq 1-\Pr[\text{some }r_i = w_i']\\
&\geq 1-\ell \frac{1}{2^n} = 1-\frac{\poly(n)}{2^n} = \ngl(n)
\end{align*}
\end{proof}
\bnote{even if all $t$ are inserted into the bad block its okay right now.  My goal is to decrease the error tolerance of the code below $t$.}
We now turn to the question of how many $0\rightarrow 1$ bits flips occur between $c, c'$.  Recall that $\dis(w, w')\leq t$ furthermore recall that $w, w'$ are selected independently of the fuzzy extractor, in particular, independently of $c$.  We break the indices of $c = c_1,..., c_\ell$ into two components those where $c_i=1$ and those $c_i = 0$.  Denote these two components by $c_{I_1}$ and $c_{I_0}$.  Our goal is to bound how many of the places where $w, w'$ are inserted into $c_{I_0}$.  Indices $i$,  where $w_i \neq w_i'$ and $c_i = 1$ will be ``corrected'' as $w_i'$ yield $c_i' = 1$ with overwhelming probably~(see \lemref{lem:no 1 to 0 flips}).  Since $w_{err}$ be set of indices where $w, w'$ differ then for all $j\in w_{err}, \Pr[j\in c_{I_0}] = 1/2$.  This is a binomial distribution with $p=1/2$ and $n=t$.  Denote by $X =|\{j | w_j \neq w_j' \wedge j\in c_{I_1}|$. Thus, $\expe[X] = t/2$.  Using Hoeffding's inequality~\cite{hoeffding1963probability}, one has that 
\begin{align*}
\Pr[X\leq t(1/2+\alpha)] \leq 1-e^{-2\alpha^2 t}
\end{align*}
This leads us to the following lemma:
\begin{lemma}
\label{lem:correctness holds}
If $t = \omega(\log n)$ and that $\dis(w, w')\leq t$ then for any constant $\alpha>0$, then $\Pr[\dis(c, c')\geq t(1/2+\alpha)] \leq 1-e^{-2\alpha t} = 1-\ngl(n)$.
\end{lemma}
\bnote{This analysis is done assuming that the two sets are the same size.  Need to worry about variation in the number of $1$s and $0$s in $c$.  Need to show that for most codes (linear?) the variance in Hamming weight is small.}
\textbf{Note:} If $t = O(\log n)$ the only error tolerance for $C$ that results in decoding with overwhelming probability is $t$.  However, in most use cases, we expect $t=\omega(\log n)$.

\subsection{Supported Sources}
\label{sec:supported sources}
\bnote{need to complete this section and redo the proof in 3.2 with this type of source.}
We now describe a set of high entropy sources appropriate for our construction.  After giving the description of the distribution, we give examples of distributions that fit in this class and provide additional characterizations.

\begin{definition}
\label{def:block guessable}
Let $\mathcal{O}_{w_1,..., w_k}$ be an oracle that return $\mathcal{O}_{w_1,..., w_k}(i, w_i')=\left( w_i\overset{?}=w_i'\right)$.
A source $W = W_1||...|W_k$ is a $(q, \alpha, \ell)$-\emph{guessable block distribution} if for any adversary $A$~(not bounded in time or space) with oracle access to $\mathcal{O}$ making at most $q$ queries there exists a set $S$ of size at least $\ell$ such that 
\[
\forall i\in S, \Hav(W_i |View(A^{\mathcal{O}_{W}(\cdot, \cdot)}))\geq \alpha.
\]
\end{definition}
This type of source seems to be inherently adaptive and is difficult to formulate in a clean information-theoretic notion.  However, we can show necessary and sufficient types of sources for a guessable block distribution.  We begin by defining the entropy jump of a block source which will appear in our characterization.

@@

\begin{claim}
Let $W$ be a $(q, \alpha, \ell)$-guessable block distribution.  Then for all orderings $i_1, ...., i_k$ there is some $j\leq k-\ell$ such that $J(i_j)> \log q + \alpha$.
\end{claim}
\begin{proof}
Suppose not, for convenience assume the ordering that violates this condition is $1,..., k$.  We describe a distribution $W$ where $\forall 1\leq j\leq k-\ell$, $J(j)\leq \log q +\alpha$. Define $W$ as follows $W_1 = c_1,...., W_{j-1} = c_{j-1}$ are constants.
\end{proof}

\subsection{Discussion}
Security of \consref{cons:first construction} relies on a large number of dimensions which are unpredictable.  Correction occurs in a small number of dimensions that are allowed to vary completely.  As described in \secref{sec:supported sources}, \consref{cons:first construction} is not secure for an arbitrary min-entropy source, it must contain a significant number of dimensions that are hard to guess by equality queries~(meaning they have min-entropy).

To show that $gap$ can be negative for \consref{cons:first construction}, we first calculate the size of the Hamming ball.  We allow $t$ errors with an alphabet of size $2^n$.  This means that
\begin{align*}
\log |B_t(\cdot)| &= \log \sum_{i=0}^t {\ell \choose i} (2^n-1)^i\\
&> \log {\ell \choose t} (2^n-1)^t\\
&=\Omega(tn) + \log {\ell\choose t}
% \log (2^n)^{H_{2^n}(t/\ell)\ell - o(\ell)} =n( H_{2^n}(t/\ell)\ell -o(\ell) )
\end{align*}
We now ask how much entropy is necessary for security?  The first presented type of source was where each block was independent and had super-logarithmic min-entropy.  This means that $\Hoo(W) \approx \ell \omega(\log n)$.  Thus in the setting where $t = \omega(\ell)$ we obtain that 
\[
gap = \Hoo(W) - \log |B_t(\cdot)| < \ell \omega(\log n) -\left( \omega(\ell n) + \log {\ell \choose t}\right)< 0. 
\]
This \consref{cons:first construction} is the first fuzzy extractor with a negative $gap$.\footnote{We can make gap smaller by recalling that note every block must contribute fresh entropy.  If each block is derived from a single source with super-logarithmic entropy then the starting entropy need only be $\omega(\log n)$.}

\textbf{Note:} If $t = o(\ell)$ then the value of $gap$ is not clear, there are some $\ell$ for which it will be negative.  However in the setting where $n>>\ell$ and a constant number of errors are being corrected $gap$ may be positive.  However, we extract from at most $\ell$ bits and thus need $\ell$ to be large enough for reasonable application security.  Setting $ \ell , t, n$ all of roughly the same size provides a natural balance of parameters.

\textbf{Limitations of \consref{cons:first construction}:}  There are three major drawbacks to our construction.  The first is that Hamming distance is over an alphabet of exponential size, which precludes use is several major settings.  The second is that we leak equality information about particular blocks, this may be sensitive information in the presence of auxiliary information.  Lastly, we obtain only a single bit from each block of $W_i$, in settings where entropy is at a premium this seems wasteful.
\section{Second construction}
\bibliographystyle{alpha}
\bibliography{crypto}
